---
title: "Simulation on the focusArea help example"
output: pdf_document
date: "2023-07-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("/home/onyxia/work/comparing_maps/outputs/simulation_on_focus_area_example_results_500sims.RData")
```


Observation: the KWD provided by the focusArea function seems not meaningful/realistic.

Conjecture: We have to divide the raw results by either the focus_area total count or 
the whole map total count. (Is there any option?)

Objective: From the example shown in the help of the package, we compare the raw results 
of the focusArea function with the raw results divided by the total count of the values 
within, respectively, the focus area and the whole map.

Method: Monte-Carlo estimation of each option of the kwd.

- "kwd" denotes the raw result of the `focusArea` function
- "kwd_fa" denotes the "kwd" estimator divided by the total count within the focus area
- "kwd_whole" denotes the "kwd" estimator divided by the total count within the whole map.


## Simulation

```{r, eval=FALSE}
library(SpatialKWD)

set.seed(40889)
n <- 500
res <- matrix(nrow = n, ncol=3)
colnames(res) <- c("kwd","kwd_fa","kwd_whole")

for(i in seq_len(n)){
  N = 90
  Xs <- as.integer(runif(N, 0, 31))
  Ys <- as.integer(runif(N, 0, 31))
  coordinates <- matrix(c(Xs, Ys), ncol=2, nrow=N)
  index_coordinates_fa <- coordinates[
    coordinates[,1] >= 10 &
      coordinates[,2] >= 10 &
      coordinates[,1] <= 20 &
      coordinates[,2] <= 20,
  ]
  # Random weights
  test1 <- matrix(runif(2*N, 0, 1), ncol=2, nrow=N)
  test1_fa <- test1[index_coordinates_fa,]
  # Compute distance
  print("Compare one-to-one with exact algorithm:")
  d <- focusArea(coordinates, Weights=test1,
                 x=15, y=15, radius=5,
                 method="exact", recode=TRUE, verbosity = "info")
  res[i,] <- c(d$distance, d$distance/sum(test1_fa[,2]), d$distance/sum(test1[,2]))
}
```

```{r}
cummeans <- apply(res, 2, cumsum)/seq_along(res[,1])
```


## Results 


### Mean Estimates

```{r}
(d_means <- colMeans(res))
```


### Range of the values obtained during the simulations

```{r}
(range_kwd <- apply(res, 2, range))
```


### Stability of the estimations

```{r}
plot(cummeans[,"kwd"], type="l")
```


```{r}
plot(cummeans[,"kwd_fa"], type="l")
```


```{r}
plot(cummeans[,"kwd_whole"], type="l")
```

## Can we conclude ?

To decide which is the good estimator of the kwd is not so obvious:  

- the raw kwd seems difficult to interpret, being constantly outside [0,1];
- the mean of each of the other two estimators are inside [0;1]
- As the values are randomly drawn, we can expect a quite large value of the KWD,
and in that case, we coul tend to choose the kwd_fa estimator.
- But only the kwd_whole has all its values within [0;1]

That's why I would choose the kwd_whole estimator.
This needs to be theorically confirmed.


