---
title: "Using SpatialKWD package"
author: "Julien Jamme"
date: "2023-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = file.path(getwd(),".."))
```


## Objectives

- Learn how to use the functions within the `SpatialKWD` package
- How the KWD measure reacts to different kinds of compared maps
- How different from a standard measure is the KWD measure when comparing maps ?

- Learn about the time computation : How big can be the maps to compare them in a realistic time  ?
- L = 2/L=3/L=10 => how does this paramter influence the quality of the measure and the computation's time ?



```{r}
if(! "SpatialKWD" %in% installed.packages()[,1]) install.packages("SpatialKWD")
library(ggplot2)
library(purrr)
library(dplyr)
library(SpatialKWD)
library(raster)

source("R/functions.R", encoding = "UTF-8")
```

```{r}
shift_weight <- function(coordinates, weights, type = "horizontal", shift = 1){
  require(dplyr)
  
  data <- cbind(coordinates[,1:2], weights[,1]) %>% 
    as_tibble()
  names(data) <- paste0("V",1:3)
  shifted <- data %>% {
    if(type == "horizontal") arrange(.,V2) %>% group_by(V2) else arrange(.,V1) %>% group_by(V1)} %>% 
    mutate(V4 = lag(V3, default = 0)) %>% {
      if(type != "horizontal") arrange(.,V2,V1) else .} %>% 
    pull(V4)
  return(cbind(weights, shifted))
}
```





## `CompareOneToOne()`

This function lets compare a distribution with one another.

### Example of the documentation

```{r}
set.seed(4081789)
# Random coordinates
N = 90
Xs <- as.integer(runif(N, 0, 31))
Ys <- as.integer(runif(N, 0, 31))
coordinates <- matrix(c(Xs, Ys), ncol=2, nrow=N)
# Random weights
test1 <- matrix(runif(2*N, 0, 1), ncol=2, nrow=N)
```


```{r}
matrix_to_maps(coordinates, test1) |> plot()
```



```{r}
d_OTO <- compareOneToOne(
  coordinates, Weights=test1, 
  L = 2,
  method="approx",recode=TRUE, 
)
```

```{r}
rmse(test1)
```


```{r}
str(d_OTO)
```

### Other simple examples

#### Parameters

```{r}
wid = 10
len = 10
N = wid * len

Xs <- 1:wid
Ys <- 1:len

coordinates <- expand.grid(Xs, Ys) |> as.matrix()
```


#### Random noise

```{r}
set.seed(4081789)
weights <- matrix(runif(2*N, 0, 1), ncol=2, nrow=N)
```


```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_rn <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_rn$distance
```

```{r}
hellinger_rn <- hellinger(weights)
hellinger_rn
```

```{r}
rmse_rn <- rmse(weights)
rmse_rn
```

#### Two identical maps


```{r}
weights <- matrix(runif(N, 0, 1), ncol=1, nrow=N)
weights <- cbind(weights, weights)
```


```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_id <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_id$distance
```

```{r}
hellinger_id <- hellinger(weights)
hellinger_id
```

```{r}
rmse_id <- rmse(weights)
rmse_id
```

#### Two horizontal shifted maps

```{r}
set.seed(4081789)
weights <- matrix(runif(N, 0, 1), ncol=1, nrow=N)
weights <- shift_weight(coordinates, weights)
```


```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_sh <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_sh$distance
```

```{r}
rmse_sh <- rmse(weights)
rmse_sh
```

#### Two vertical shifted maps

```{r}
set.seed(4081789)
weights <- matrix(runif(N, 0, 1), ncol=1, nrow=N)
weights <- shift_weight(coordinates, weights, type = "vertical")
```


```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_sv <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_sv$distance
```

```{r}
hellinger_sv <- hellinger(weights)
hellinger_sv
```

```{r}
rmse_sv <- rmse(weights)
rmse_sv
```


#### Chessboard maps

```{r}
vs <- c(0.25,0.75)
b1 <- matrix(vs[1], ncol = len/2, nrow = wid/2)
b2 <- matrix(vs[2], ncol = len/2, nrow = wid/2)
w1 <- rbind(cbind(b1,b2),cbind(b2,b1))
w2 <- rbind(cbind(b2,b1),cbind(b1,b2))

weights <- cbind(
  from_matrix_to_column(w1),
  from_matrix_to_column(w2)
)
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_ch <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_ch$distance
```

```{r}
hellinger_ch <- hellinger(weights)
hellinger_ch
```

```{r}
rmse_ch <- rmse(weights)
rmse_ch
```

#### Chessboard maps 2

```{r}
vs <- c(0.25,0.75)
b1 <- matrix(vs[1], ncol = 1, nrow = 1)
b2 <- matrix(vs[2], ncol = 1, nrow = 1)

w1 <- matrix(rep(c(b1, b2), 50), nrow=9, ncol=9, byrow = TRUE)
w2 <- matrix(rep(c(b2, b1), 50), nrow=9, ncol=9, byrow = TRUE)


weights <- cbind(
  from_matrix_to_column(w1),
  from_matrix_to_column(w2)
)

coordinates <- as.matrix(expand.grid(X=1:9, Y=1:9))
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_ch2 <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_ch2$distance
```

```{r}
hellinger_ch2 <- hellinger(weights)
hellinger_ch2
```

```{r}
rmse_ch2 <- rmse(weights)
rmse_ch2
```


## Sensitivity to NA/empty values

```{r}
vs <- c(NA,0.75)
b1 <- matrix(vs[1], ncol = 1, nrow = 1)
b2 <- matrix(vs[2], ncol = 1, nrow = 1)

w1 <- matrix(rep(c(b1, b2), 50), nrow=9, ncol=9, byrow = TRUE)
w2 <- matrix(rep(c(b2, b1), 50), nrow=9, ncol=9, byrow = TRUE)


weights <- cbind(
  from_matrix_to_column(w1),
  from_matrix_to_column(w2)
)

coordinates <- as.matrix(expand.grid(X=1:9, Y=1:9))
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_na <- compareOneToOne(coordinates, weights)
kwd_na$distance
```

Distance is null. The presence of NA values seems to perturb the computation of the distance.

### NA to zeros

```{r}
weights[is.na(weights)] <- 0
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_zeros <- compareOneToOne(coordinates, weights, verbosity = "silent")
kwd_zeros$distance
```

## Work on autocorrelated data: the Meuse data

```{r}
data(meuse.grid, package = "sp")
dim(meuse.grid)
```


### Sensitivity to injection of noise in the data

```{r}
set.seed(4081789)
coordinates <- as.matrix(meuse.grid[,1:2])
weights <- cbind(
  meuse.grid[,"dist"] + 10,
  meuse.grid[,"dist"] + 10 + rnorm(nrow(meuse.grid), 0, 0.001),
  meuse.grid[,"dist"] + 10 + rnorm(nrow(meuse.grid), 0, 0.01),
  meuse.grid[,"dist"] + 10 + rnorm(nrow(meuse.grid), 0, 0.1),
  meuse.grid[,"dist"] + 10 + rnorm(nrow(meuse.grid), 0, 0.5),
  meuse.grid[,"dist"] + 10 + rnorm(nrow(meuse.grid), 0, 1)
)
```

Remark: the +10 operation is to ensure the values remains positive.

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_autocorr <- compareOneToMany(coordinates, weights, recode = TRUE)
kwd_autocorr
```

```{r}
hellinger_autocorr <- hellinger(weights)
hellinger_autocorr
```

```{r}
rmse_autocorr <- rmse(weights)
rmse_autocorr
```


```{r}
tibble(
  sd = c(0.001,0.01,0.1, 0.5, 1),
  kwd = kwd_autocorr$distance,
  rmse = rmse_autocorr
) %>%
  tidyr::pivot_longer( cols = c("kwd", "rmse"), names_to = "measure", values_to = "val") %>% 
  ggplot(aes(x=sd, y = val, color = measure)) +
  geom_line() +
  geom_point() +
  scale_x_log10("sd (log-10 scale)") +
  theme_minimal()
```


### Sensitivity to the magnitude

We compare two maps different only on the magnitude of the cell values: map2 is 
like map1 + A or like map1*B.

```{r}
summary(meuse.grid[,"dist"])
```

```{r}
mean_dist <- mean(meuse.grid[,"dist"])
```

#### translation of the distribution

Map2 = Map1 + A

```{r}
coordinates <- as.matrix(meuse.grid[,1:2])
weights <- purrr::map(
  c(0,1,2,10,100,1000),
  function(a)  meuse.grid[,"dist"] + a*mean_dist
) %>% purrr::reduce(cbind)
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_magn <- compareOneToMany(coordinates, weights)
kwd_magn$distance
```

```{r}
hellinger_magn <- hellinger(weights)
hellinger_magn
```

```{r}
rmse_magn <- rmse(weights)
rmse_magn
```

#### Homothetic transformation of the distribution

Map2 = Map1 * B


```{r}
coordinates <- as.matrix(meuse.grid[,1:2])
weights <- purrr::map(
  c(1/mean_dist,1,2,10,100,1000),
  function(a)  meuse.grid[,"dist"] * a*mean_dist
) %>% purrr::reduce(cbind)
```

```{r}
matrix_to_maps(coordinates, weights) |> plot()
```

```{r}
kwd_hom <- compareOneToMany(coordinates, weights)
kwd_hom$distance
```

```{r}
hellinger_hom <- hellinger(weights)
hellinger_hom
```

```{r}
rmse_hom <- rmse(weights)
rmse_hom
```














